Delayed dichromatism as a convenient tool to disentangle the effects of survival and productivity on the population dynamics in waterfowl

### 2024-06-04 Code/Draft_ruddy_duck_vital_rates.Rmd
layout fix
Revision before submission to bioarxiv

---
lang: en-EN
bibliography: "`r paste0('C:/Users/', Sys.info()[['user']], '/Dropbox/JabRef/Library_OFB.bib')`"
csl: https://www.zotero.org/styles/apa # https://osf.io/download/6hw3q/ 
link-citations: true
title: "Delayed dichromatism as a convenient tool to disentangle the effects of survival and productivity on the population dynamics in waterfowl"
author: 
- Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
- Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
- Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
- Matthieu Guillemain^[Office Français de la Biodiversité, France - matthieu.guillemain@ofb.gouv.fr]
- Jean-François Maillard^[Office Français de la Biodiversité, France - jean-francois.maillard@ofb.gouv.fr]
- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%d %B, %Y')`"
abstract: |
  Monitoring the number of individuals is by far the most popular strategy for studying the environmental factors that determine population dynamics and for measuring the effectiveness of management actions aimed at population recovery, control, or eradication. unfortunately, monitoring numbers is inefficient in identifying the mechanisms underlying demographic processes and, in particular, in assessing the extent to which population growth rate is influenced by changes in adult survival rather than variations in reproductive parameters. The usual methods for determining adult survival and productivity, such as analysing capture-mark-recapture data and game-hunting bag data, rely on capturing animals, which has three major drawbacks. Firstly, the selectivity of capture methods can lead to a distorted representation of the underlying population structure. Secondly, the capture and release of animals is quite invasive, causing stress and posing an additional threat to the often already endangered species. Thirdly, release may simply be impossible, especially for invasive alien species that represent a threat to native taxa. In many waterfowl species, sexual dichromatism is observed in adults, while immatures of both sexes display a cover-up plumage similar to that of adult females. Using two non-native populations of Ruddy duck, a species that exhibits this pattern, we present a non-invasive method to estimate the respective contribution of adult survival and productivity (also called recruitment rate) to the population growth rate. We exemplify how changes in the apparent proportion of males in counts (due to female-like immature males gradually acquiring adult male plumage) can be exploited to split the population growth rate into adult survival and recruitment rates, which contributes significantly to understanding the demographic impact of two different eradication strategies.

  **Keywords**: age ratio - alien species - demography - fecundity - harvesting - intrinsic growth rate - Oxyura jamaicensis - reproductive success - sex ratio - vital rate
output: 
    bookdown::pdf_document2:
      latex_engine: xelatex
      toc: false
      keep_md: false
      keep_tex: false
    bookdown::word_document2:
      toc: false
header-includes:
  - \usepackage{float} # allows to hold figure position
  - \usepackage[left]{lineno} # line numbering
---

@@

Assessing the growth rate of populations is a first step towards a better understanding of the factors underlying the dynamics of natural populations [e.g. @Niel2005]. It is also crucial for measuring the effectiveness of management actions that are taken to recover, control, or eradicate populations [@Shea1998]. Among the approaches available to managers to achieve these goals, those that rely on monitoring the number of individuals (counts) are by far the most popular [@Rintala2022]. In many cases, these methods allow the investigation of environmental factors underlying changes in population size, and thus help in the implementation of mitigating actions [@Faillettaz2019]. Unfortunately, relying solely on the monitoring of population abundance generally hinders the fundamental understanding of the demographic mechanisms underlying changes in population growth rates. More specifically, on the basis of counts alone, it is almost impossible to assess the extent to which population growth rate is affected by changes in adult survival rather than to variations in productivity [@Austin2000], which is defined as the product of all reproductive parameters. To assess the relative sensitivity of population growth rate to factors affecting adult survival or reproductive parameters, demographers usually rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992]. In practise, however, capture and release of individuals cannot always be relied upon, as the legal status of the species precludes any additional disturbance (e.g. critically endangered species) or prohibits the release of live individuals (e.g. invasive species or pests). Even though capture-mark-recapture methods are efficient for assessing demographic parameters, they have potential drawbacks, such as their invasiveness, which can affect behaviour and thus survival or reproductive success, and their affordability when time and money are scarce. Genetic monitoring is a non-invasive alternative to capture-mark-recapture, but it is costly and requires large samples [@Caniglia2011].

Most of the time, therefore, managers make the best of a bad job by using counts as a viable option for tracking changes in population size and then population growth rate to assess the relevance of management actions. Distinguishing between immatures and adults in counts allows estimation of adult survival rate, i.e. the proportion of breeders that have survived for one year, and productivity/recruitment rate, i.e. the number of immatures produced per breeder that reach sexual maturity, which are, by definition, recruited in the breeding population. This additional effort makes it possible to assess the relative influence of each of these components on the population growth rate. Alternatively, the assessment of age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of game species, including ducks and geese, but suffers from intractable biases [@Fox2018]. Of course, the latter approach is not appropriate for protected/endangered species.

Here we exemplify how, in dimorphic species, delayed sexual maturity of males can be used to estimate adult survival and recruitment rates by distinguishing male-like and female-like individuals in repeated winter counts. We used two non-native European populations of Ruddy duck (*Oxyura jamaicensis*) as a study model. Like many other duck species, Ruddy duck is dimorphic: newborn males look like females until the prenuptial moult, which takes place in late winter. As a result of the late prenuptial moult of immatures, the apparent proportion of males increases over the course of the wintering season. These changes in the apparent proportion of males during this period are therefore directly related to the proportion of immatures in the populations and thus to the reproductive success of the previous breeding season. The "apparent sex ratio" method developed in this study was used to estimate adult survival and recruitment rates and evaluate the effects of two different eradication strategies used in Great Britain and France, respectively.

@@

The trends and status of the White-headed duck and Ruddy duck populations used to be determined exclusively by counts. It was therefore not possible to assess the factors impairing the recovery of the former and the effectiveness of the eradication programme for the latter. In particular, counts alone cannot be used to assess the relative impact of changes in adult survival and recruitment rates on population growth rates, which is a prerequisite for identifying limiting factors.

```{r count, fig.cap = "(ref:count)"}
knitr::include_graphics("../Output/plot_paper_2_saved.png")
```

@@

Interestingly, delayed dichromatism allows the estimation of adult survival and recruitment/productivity by observing the seasonal evolution of the apparent sex ratio and making few and reasonable assumptions (see next section).

@@

```{r flock, fig.cap = "(ref:flock)"}
knitr::include_graphics("../Output/flock.png")
```

```{r pmal, fig.cap = "(ref:pmal)"}
knitr::include_graphics("../Output/plot_paper_1_saved.png")
```

We monitored the demography of Ruddy duck populations in both Great Britain (GB) and France (FR) by conducting extensive counts on the wintering grounds between 1 December and 31 January. In addition, the apparent sex ratio could be determined using winter counts that distinguished between female-like and male-like individuals. Such counts were carried out in Great Britain from 2006 to 2012, which corresponded to a period of sharp population decline (Figure \@ref(fig:count)). In France, count surveys taking into account apparent sex were conducted in 1999, 2001-2009, and 2012-2019. The French population grew rapidly in the first years of monitoring and then stabilised from 2006 onwards as a result of high culling pressure. In both countries, culling was carried out under strict official control so that the exact number of individuals culled was known. Although the age ratio of birds culled before 2009 could not be determined with certainty in France, it was still possible to roughly categorise the culling pressure in both countries into three categories: "no culling" before 1999, "low culling" between 1999 and 2005 and "high culling" from 2006 onwards (see Figure \@ref(fig:sup) in Supplementary Material).

@@

The proportion of immature males is the difference between the proportion of female-like individuals between winter and spring. The proportion of all immature individuals is twice as high if we assume an even sex ratio at birth [@Bellrose1961; @Blums1996] and comparable mortality rates between immature females and immature males [@Ramula2018; @Wood2021]. However, spring counts suffer from intractable problems, such as the different detectability of females and males during this season. Fortunately, the proportion of immatures can also be estimated from winter counts alone, assuming a known and unchanged adult sex ratio along the monitoring period [@Devineau2010].

The adult sex ratio is therefore a prerequisite for estimating the proportion of immatures. This ratio could be derived from culling data as we determined the sex of each adult culled as part of the eradication programmes in both Great Britain and France. Nevertheless, the low number of individuals culled in France prevented from getting robust adult sex ratio. In Great Britain, adult sex ratios did not differ from month to month, so data from culled adults collected throughout the whole biological cycle were pooled together each year. The proportion of males also did not differ statistically between years in which more than 500 adults were culled. This is consistent with previous findings showing that the adult sex ratio in ducks is usually very stable in the short term [@Wood2021] but can fluctuate in the long term. Therefore, we pooled the data from all adults culled in the control programme to estimate the adult sex ratio, see Equation \@ref(eq:pm).

@@

| Variable name         | Description                                                                                                |
|:-|:-------|
| *Observation (input)* |                                                                                                            |
| $AM$                  | Total number of adult males culled as part of the eradication programme                                      |
| $AF$                  | Total number of adult females culled as part of the eradication programme                                    |
| $CML_{i, t}$          | Number of male-like individuals counted in population $i$ in year $t$                                      |
| $C_{i, t}$            | Number of individuals counted in population $i$ in year $t$                                                |
| $N_{i, t}$            | Size of the population $i$ in year $t$ (maximum number of individuals counted in the wintering population) |
| *Parameter (output)*  |                                                                                                            |
| $p(m|a)$              | Proportion of males among adults (or probability of being a male knowing it is an adult)                      |
| $p(a \cap m)_{i, t}$  | Proportion of adult males in population $i$ in year $t$                                                    |
| $p(a)_{i, t}$         | Proportion of adults in population $i$ in year $t$                                                         |
| $p(re)_{i, t}$        | Proportion of immatures/recruits in population $i$ in year $t$                                             |
| $s_{i, t}$            | Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$                  |
| $r_{i, t}$            | Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per individual in year $t - 1$     |
| $S_{i, t}$            | Number of adults in population $i$ in year $t$                                                             |
| $R_{i, t}$            | Number of recruits in population $i$ in year $t$                                                           |

```{=tex}
\begin{equation}
p(m|a) \sim {\sf Beta}(AM, AF)
(\#eq:pm)
\end{equation}
```
```{=tex}
\begin{align}
CML_{i, t} & \sim {\sf Binom}(p(a \cap m)_{i, t}, C_{i, t}) \notag \\
& \sim {\sf Binom}(p(m|a).p(a)_{i, t}, C_{i, t}) \notag \\
& \sim {\sf Binom}(p(m|a).(1 - p(re)_{i, t}), C_{i, t})
(\#eq:pic)
\end{align}
```
```{=tex}
\begin{equation}
\begin{aligned}
s_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.(1 - p(re)_{i, t}) \\
r_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.p(re)_{i, t}
\end{aligned}
(\#eq:vr)
\end{equation}
```
```{=tex}
\begin{equation}
\begin{aligned}
S_{i, t} & = N_{i, t}.(1 - p(re)_{i, t}) \\
R_{i, t} & = N_{i, t}.p(re)_{i, t}
\end{aligned}
(\#eq:VR)
\end{equation}
```
```{r outline, fig.cap = "(ref:outline)"}
knitr::include_graphics("../Output/outline.png")
```

@@

The relevance of the method was assessed by testing the likelihood of the estimates of both vital rates. For the adult survival rate, we checked whether the values were contained in the interval $[0;1]$, and compared our estimates with those in the literature [e.g. @Buxton2004; @Krementz1997; @Nichols1992, @Nichols1997]. Validation of the recruitment rate was more difficult because if the lower limit is 0, the upper limit is a product of all maximum values of its individual components, i.e. nesting rate, nesting success, clutch-size, hatching success, pre-fledging survival, and fledging survival. These reproductive parameters may be known for some duck species, but the last one is never mentioned [e.g. @Baldassarre2014].

@@

| Variable name | Description                                                                                 |
|:-|:---------|
| $N_{t}$       | Number of adults in year $t$                                                                |
| $D_{t}$       | Number of adults dead during year $t$                                                       |
| $R_{t}$       | Number of recruits in year $t$                                                              |
| $s_{t}$       | Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$   |
| $r_{t}$       | Recruitment rate, i.e. number of recruits in year $t$ produced per individual in year $t-1$ |
| $\lambda_{t}$ | Growth rate of the population between year $t-1$ and year $t$                               |

```{=tex}
\begin{align} 
N_{t} & = N_{t-1} - D_{t-1} + R_{t} \notag \\
N_{t} & = N_{t-1} - (1 - s_{t}).N_{t-1} + r_{t}.N_{t-1} \notag \\
N_{t} & = s_{t}.N_{t-1} + r_{t}.N_{t-1} \notag \\
\frac{N_{t}}{N_{t-1}} & = s_{t} + r_{t} \notag \\
\lambda_{t} & = s_{t} + r_{t} \notag \\
r_{t} & = \lambda_{t} - s_{t}  
(\#eq:GR)
\end{align}
```
The maximum adult survival rate was defined as a uniform distribution on the interval $[0.7;1]$. The lower limit corresponded to the upper range of survival rates observed in waterfowl species of similar weight, and the upper limit was set to 1, as survival rates of long-lived waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992; @Koons2014].

For both populations, the maximum growth rates occurred during their geographic expansion phase, which preceded the start of the eradication programmes (Figure \@ref(fig:count)). To estimate robust maximum population growth rates for both populations, we smoothed the annual growth rates over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L). We discarded the data for Great Britain before 1972 as the estimates were likely noisy when the population was low (below 50 individuals) (Figure \@ref(fig:count)). After reaching the threshold of 1,000 individuals, the population growth in Great Britain (GB) showed a strong inflexion, although culling had not yet started (Figure \@ref(fig:count)). This observation suggests that a negative density-dependent process could occur beyond 1,000 individuals and led us to consider only the first sequence of the time series to infer the maximum growth rate in Great Britain, i.e. 1972-1981. For the French (FR) population, the sequence without culling pressure covered the period 1994-1999.

| Variable name    | Class     | Description                                                                                              |
|:-|:-|:--------|
| $N0_{i, J}$      | Parameter | Intercept of the regression model                                                                        |
| $\lambda_{i, J}$ | Parameter | Average population growth rate over a restricted time interval $J$ for a population $i$ (in $year^{-1}$) |
| $t \in J$        | Index     | Year index within the time interval $J$                                                                  |
| $\sigma_{i, J}$  | Parameter | Standard deviation of the regression model                                                               |

```{=tex}
\begin{equation}
log(N_{i, t}) \sim {\sf Norm}(N0_{i, J} + log(\lambda_{i, J}).t, \sigma_{i, J})
(\#eq:L)
\end{equation}
```
Estimating vital rates within an acceptable range, as defined above, would not prove that our modelling approach correctly reflects the actual interannual variability. To determine this, we compared the proportions of immatures and the vital rates derived from counts with those derived from culling, i.e. from individuals culled as part of the eradication programme. As we restricted the dataset to years with more than 100 individuals culled in winter for sake of precision, the analysis only covered five years of the GB population time series. The proportions of immatures were estimated from the culled individuals by checking the presence of the bursa of Fabricius, which is only present in immature individuals [@Hochbaum1942]; see Equation \@ref(eq:pis). Adult survival and recruitment rates were then derived from the proportion of immatures and the Equation \@ref(eq:vr).

| Variable name | Class        | Description                                                              |
|:-|:-|:--------|
| $SI_{i, t}$   | Observation  | Number of immatures culled in the wintering population $i$ in year $t$   |
| $S_{i, t}$    | Observation  | Number of individuals culled in the wintering population $i$ in year $t$ |

```{=tex}
\begin{equation}
SI_{i, t} \sim {\sf Binom}(p(re)_{i, t}, S_{i, t})
(\#eq:pis)
\end{equation}
```
## Assessing the effects of culling strategies

@@

We used the Bayesian framework to implement all submodels presented in Materials & methods section. The Bayesian framework is both straightforward and efficient in propagating error through the parameters. We used uninformative priors for all parameters. Since the maximum growth rate is a life history trait expected to be stable between populations of the same species, we used an uninformative hierarchical prior for this parameter. We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, and discarded the first 2,000 samples as burn-in. We checked the chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992). The models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022]. The values **X[Y; Z]** reported in Results section are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions. We preferred the median to the mean because it is more robust to skewed distributions.

@@

```{r vital, fig.cap = "(ref:vital)"}
knitr::include_graphics("../Output/plot_paper_3.png")
```

@@

When there was no culling pressure, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), namely 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for GB and FR, respectively. These values corresponded to an increase in the population of 45% and 52% per year, respectively. Assuming a maximum adult survival rate of 0.85 [0.7; 1] for both populations, we derived maximum recruitment rates of 0.60 [0.42; 0.78] and 0.68 [0.36; 1.06] for GB and FR, respectively.

>>>>

---
lang: en-EN
bibliography: "`r here::here('Data', 'vital_rates.bib')`"
csl: https://www.zotero.org/styles/apa # https://osf.io/download/6hw3q/ 
link-citations: true
title: "Delayed dichromatism as a convenient tool to disentangle the effects of survival and productivity on the population dynamics in waterfowl"
author: 
- Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
- Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
- Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
- Matthieu Guillemain^[Office Français de la Biodiversité, France - matthieu.guillemain@ofb.gouv.fr]
- Jean-François Maillard^[Office Français de la Biodiversité, France - jean-francois.maillard@ofb.gouv.fr]
- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%d %B, %Y')`"
abstract: |
  | Monitoring the number of individuals is by far the most popular strategy for studying the environmental factors that determine population dynamics and for measuring the effectiveness of management actions aimed at population recovery, control, or eradication. unfortunately, monitoring numbers is inefficient in identifying the mechanisms underlying demographic processes and, in particular, in assessing the extent to which population growth rate is influenced by changes in adult survival rather than variations in reproductive parameters. The usual method for determining adult survival and productivity relies on capturing and marking, which has three major drawbacks. Firstly, the selectivity of capture methods can lead to a distorted representation of the underlying population structure. Secondly, the capture and release of animals is quite invasive, causing stress and posing an additional threat to the often already endangered species. Thirdly, release may simply be impossible, especially for invasive alien species that represent a threat to native taxa. In many waterfowl species, sexual dichromatism is observed in adults, while immatures of both sexes display a plumage pattern similar to that of adult females. Using two non-native populations of Ruddy duck, a species that exhibits this plumage pattern, we present a non-invasive method to estimate the respective contribution of adult survival and productivity (also called recruitment rate) to the population growth rate. We exemplify how changes in the apparent proportion of males in counts (due to female-like immature males gradually acquiring adult male plumage) can be exploited to split the population growth rate into adult survival and recruitment rates, which contributes significantly to understanding the demographic impact of two different eradication strategies.
  | 
  | **Keywords**: age ratio - alien species - demography - fecundity - harvesting - intrinsic growth rate - Oxyura jamaicensis - reproductive success - sex ratio - vital rate
linestretch: 2
output: 
    bookdown::pdf_document2:
      latex_engine: xelatex
      toc: false
      keep_md: false
      keep_tex: false
    bookdown::word_document2:
      toc: false
header-includes:
  - \usepackage{float} # allows to hold figure position
  - \usepackage[left]{lineno} # line numbering
---

@@

Assessing the growth rate of populations is a first step towards a better understanding of the factors underlying the dynamics of natural populations [e.g. @Niel2005]. It is also crucial for measuring the effectiveness of management actions that are taken to restore, control, or eradicate populations [@Shea1998]. Among the approaches available to managers to achieve these goals, those that rely on monitoring the number of individuals (counts) are by far the most popular [@Rintala2022]. In many cases, these methods allow the investigation of environmental factors underlying changes in population size, and thus help in the implementation of mitigating actions [@Faillettaz2019]. Unfortunately, relying solely on the monitoring of population abundance generally hinders the fundamental understanding of the demographic mechanisms underlying changes in population growth rates. More specifically, on the basis of counts alone, it is almost impossible to assess the extent to which population growth rate is affected by changes in adult survival rather than to variations in young production [@Austin2000]. To assess the relative sensitivity of population growth rate to factors affecting adult survival or reproductive parameters, demographers usually rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992]. In practise, however, capture and release of individuals cannot always be relied upon, as the legal status of the species precludes any additional disturbance (e.g. critically endangered species) or prohibits the release of live individuals (e.g. invasive species or pests). Even though capture-mark-recapture methods are efficient for assessing demographic parameters, they have potential drawbacks, such as their invasiveness, which can affect behaviour and thus survival or reproductive success, and their affordability when time and money are scarce. Genetic monitoring is a non-invasive alternative to capture-mark-recapture, but it is costly and requires large samples [@Caniglia2011].

Most of the time, therefore, managers make the best of a bad job by using counts as a viable option for tracking changes in population size and then population growth rate to assess the relevance of management actions. Distinguishing between immatures and adults in counts allows estimation of adult survival rate, i.e. the proportion of breeders that have survived for one year, and productivity/recruitment rate, i.e. the number of immatures produced per breeder that reach sexual maturity, which are, by definition, recruited in the breeding population [@Smith2001]. This additional effort makes it possible to assess the relative influence of each of these components on the population growth rate. Alternatively, the assessment of age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of game species, including ducks and geese, but suffers from intractable biases [@Fox2018]. Of course, the latter approach is not appropriate for protected/endangered species.

Here we exemplify how, in dimorphic species, delayed sexual maturity of males can be used to estimate adult survival and recruitment rates by distinguishing male-like and female-like individuals in winter counts. We used two non-native European populations of Ruddy duck (*Oxyura jamaicensis*) as a study model. Like many other duck species, Ruddy duck is dimorphic: newborn males look like females until the prenuptial moult, which takes place in late winter. As a result of the late prenuptial moult of immatures, the apparent proportions of males increases over the course of the wintering season. These changes in the apparent proportions of males during this period are therefore directly related to the proportion of immatures in the populations and thus to the reproductive success of the previous breeding season. The "apparent sex ratio" method developed in this study was used to estimate adult survival and recruitment rates and evaluate the effects of two different eradication strategies used in Great Britain and France, respectively.

@@

The trends and status of the White-headed duck and Ruddy duck populations used to be determined exclusively by counts. It was therefore difficult to assess the factors impairing the recovery of the former and the effectiveness of the eradication programme for the latter. In particular, counts alone cannot be used to assess the relative impact of changes in adult survival and recruitment rates on population growth rates, which is a prerequisite for identifying limiting factors.

\setstretch{1}
```{r count, fig.cap = "(ref:count)"}
knitr::include_graphics("../Output/plot_paper_2_saved.png")
```
\setstretch{2}

@@

Interestingly, delayed dichromatism allows the estimation of adult survival and recruitment/productivity by monitoring the seasonal evolution of the apparent sex ratio and making few and reasonable assumptions (see next section).

@@

\setstretch{1}
```{r flock, fig.cap = "(ref:flock)"}
knitr::include_graphics("../Output/flock.png")
```

```{r pmal, fig.cap = "(ref:pmal)"}
knitr::include_graphics("../Output/plot_paper_1_saved.png")
```
\setstretch{2}

We monitored the demography of Ruddy duck populations in both Great Britain (GB) and France (FR) by conducting extensive counts on the wintering grounds between 1 December and 31 January. In addition, the apparent sex ratio could be determined using winter counts that distinguished between female-like and male-like individuals. Such counts were carried out in Great Britain from 2006 to 2012, which corresponded to a period of sharp population decline (Figure \@ref(fig:count)). In France, count surveys taking into account apparent sex were conducted in 1999, 2001-2009, and 2012-2019. The French population grew rapidly in the first years of monitoring and then stabilised from 2006 onwards as a result of high culling pressure. In both countries, culling was carried out under strict official control so that the exact number of individuals culled was known. Although the age ratio of birds culled before 2009 could not be determined with certainty in France, it was still possible to roughly categorise the culling pressure in both countries into three categories: "no culling" before 1999, "low culling" between 1999 and 2005 and "high culling" from 2006 onwards (see Figure \@ref(fig:sup) in Appendix section).

@@

Our method exploits the difference between the sex ratio estimated from counts before young males acquire their male-like plumage pattern (and are therefore mistaken for females) and the sex ratio among adults (which is stable over time), which can be estimated from either spring or summer counts or from hunting bags. In the present study, we estimated the adult sex-ratio from culled individuals using the presence/absence of the bursa of Fabricius as a criterion for distinguishing adults from immatures (see hereafter). In France, it was not possible to obtain sufficiently reliable estimates of the sex ratio due to the small sample size. In Great Britain, adult sex ratios did not differ from month to month, so data from culled adults collected throughout the whole biological cycle were pooled together each year. The proportion of males also did not differ statistically between years in which more than 500 adults were culled. This is consistent with previous findings showing that the adult sex ratio in ducks is usually very stable in the short term [@Wood2021] but can fluctuate in the long term. Therefore, we pooled the data from all adults culled in the control programme to estimate the adult sex ratio, see Equation \@ref(eq:pm).

@@

\setstretch{1}
| Variable name         | Description                                                                                                |
|:-|:-------|
| *Observation (input)* |                                                                                                            |
| $AM$                  | Total number of adult males culled as part of the eradication programme                                    |
| $AF$                  | Total number of adult females culled as part of the eradication programme                                  |
| $CML_{i, t}$          | Number of male-like individuals counted in population $i$ in year $t$                                      |
| $C_{i, t}$            | Number of individuals counted in population $i$ in year $t$                                                |
| $N_{i, t}$            | Size of the population $i$ in year $t$ (maximum number of individuals counted in the wintering population) |
| *Parameter (output)*  |                                                                                                            |
| $p(m|a)$              | Proportion of males among adults (or probability of being a male knowing it is an adult)                   |
| $p(a \cap m)_{i, t}$  | Proportion of adult males in population $i$ in year $t$                                                    |
| $p(a)_{i, t}$         | Proportion of adults in population $i$ in year $t$                                                         |
| $p(re)_{i, t}$        | Proportion of immatures/recruits in population $i$ in year $t$                                             |
| $s_{i, t}$            | Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$                  |
| $r_{i, t}$            | Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per adult in year $t - 1$     |
| $S_{i, t}$            | Number of adults in population $i$ in year $t$                                                             |
| $R_{i, t}$            | Number of recruits in population $i$ in year $t$                                                           |

```{=tex}
\begin{equation}
p(m|a) \sim {\sf Beta}(AM, AF)
(\#eq:pm)
\end{equation}
```
```{=tex}
\begin{align}
CML_{i, t} & \sim {\sf Binom}(p(a \cap m)_{i, t}, C_{i, t}) \notag \\
& \sim {\sf Binom}(p(m|a).p(a)_{i, t}, C_{i, t}) \notag \\
& \sim {\sf Binom}(p(m|a).(1 - p(re)_{i, t}), C_{i, t})
(\#eq:pic)
\end{align}
```
```{=tex}
\begin{equation}
\begin{aligned}
s_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.(1 - p(re)_{i, t}) \\
r_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.p(re)_{i, t}
\end{aligned}
(\#eq:vr)
\end{equation}
```
```{=tex}
\begin{equation}
\begin{aligned}
S_{i, t} & = N_{i, t}.(1 - p(re)_{i, t}) \\
R_{i, t} & = N_{i, t}.p(re)_{i, t}
\end{aligned}
(\#eq:VR)
\end{equation}
```
```{r outline, fig.cap = "(ref:outline)"}
knitr::include_graphics("../Output/outline.png")
```
\setstretch{2}

@@

The relevance of the method was assessed by testing the likelihood of the estimates of both vital rates. For the adult survival rate, we checked whether the values were contained in the interval $[0;1]$, and compared our estimates with those in the literature [e.g. @Buxton2004; @Krementz1997; @Nichols1992, @Nichols1997]. Validation of the recruitment rate was more difficult because if the lower limit is 0, the upper limit is a product of all maximum values of its individual components, i.e. nesting rate, nesting success, clutch-size, hatching success, pre-fledging survival, and fledging survival. Although some of these reproductive parameters may be known for some duck species, the last one is never mentioned [e.g. @Baldassarre2014].

@@

\setstretch{1}
| Variable name | Description                                                                                 |
|:-|:---------|
| $N_{t}$       | Number of adults in year $t$                                                                |
| $D_{t}$       | Number of adults dead during year $t$                                                       |
| $R_{t}$       | Number of recruits in year $t$                                                              |
| $s_{t}$       | Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$   |
| $r_{t}$       | Recruitment rate, i.e. number of recruits in year $t$ produced per individual in year $t-1$ |
| $\lambda_{t}$ | Growth rate of the population between year $t-1$ and year $t$                               |

```{=tex}
\begin{align} 
N_{t} & = N_{t-1} - D_{t-1} + R_{t} \notag \\
N_{t} & = N_{t-1} - (1 - s_{t}).N_{t-1} + r_{t}.N_{t-1} \notag \\
N_{t} & = s_{t}.N_{t-1} + r_{t}.N_{t-1} \notag \\
\frac{N_{t}}{N_{t-1}} & = s_{t} + r_{t} \notag \\
\lambda_{t} & = s_{t} + r_{t} \notag \\
r_{t} & = \lambda_{t} - s_{t}  
(\#eq:GR)
\end{align}
```
\setstretch{2}

The maximum adult survival rate was defined as a uniform distribution on the interval $[0.7;1]$. The lower limit corresponded to the upper range of survival rates observed in waterfowl species of similar weight, and the upper limit was set to 1, as survival rates of long-lived waterfowl species can be very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992; @Koons2014].

For both populations, the maximum growth rates occurred during their geographic expansion phase, which preceded the start of the eradication programmes (Figure \@ref(fig:count)). To estimate robust maximum population growth rates for both populations, we smoothed the annual growth rates over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L). We discarded the data for Great Britain before 1972 as the estimates were likely noisy when the population was low (below 50 individuals) (Figure \@ref(fig:count)). After reaching the threshold of 1,000 individuals, the population growth in Great Britain (GB) showed a strong inflexion, although culling had not yet started (Figure \@ref(fig:count)). This observation suggests that beyond 1,000 individuals, a negative density-dependent process probably took place and led us to consider only the first sequence of the time series to infer the maximum growth rate in Great Britain, i.e. 1972-1981. For the French (FR) population, the sequence without culling pressure covered the period 1994-1999.

\setstretch{1}
| Variable name    | Class     | Description                                                                                              |
|:-|:-|:--------|
| $N0_{i, J}$      | Parameter | Intercept of the regression model                                                                        |
| $\lambda_{i, J}$ | Parameter | Average population growth rate over a restricted time interval $J$ for a population $i$ (in $year^{-1}$) |
| $t \in J$        | Index     | Year index within the time interval $J$                                                                  |
| $\sigma_{i, J}$  | Parameter | Standard deviation of the regression model                                                               |

```{=tex}
\begin{equation}
log(N_{i, t}) \sim {\sf Norm}(N0_{i, J} + log(\lambda_{i, J}).t, \sigma_{i, J})
(\#eq:L)
\end{equation}
```
\setstretch{2}

Estimating vital rates within an acceptable range, as defined above, would not prove that our modelling approach correctly reflects the actual interannual variability. To determine this, we compared the proportions of immatures and the vital rates derived from counts with those derived from culling, i.e. from individuals culled as part of the eradication programme. As we restricted the dataset to years with more than 100 individuals culled in winter for sake of precision, the analysis only covered five years of the GB population time series. The proportions of immatures were estimated from the culled individuals by checking the presence of the bursa of Fabricius, which is only present in immature individuals [@Hochbaum1942]; see Equation \@ref(eq:pis). Adult survival and recruitment rates were then derived from the proportion of immatures and the Equation \@ref(eq:vr).

\setstretch{1}
| Variable name | Class        | Description                                                              |
|:-|:-|:--------|
| $SI_{i, t}$   | Observation  | Number of immatures culled in the wintering population $i$ in year $t$   |
| $S_{i, t}$    | Observation  | Number of individuals culled in the wintering population $i$ in year $t$ |

```{=tex}
\begin{equation}
SI_{i, t} \sim {\sf Binom}(p(re)_{i, t}, S_{i, t})
(\#eq:pis)
\end{equation}
```
\setstretch{2}

## Assessing the effects of culling strategies

@@

We used the Bayesian framework to implement all submodels presented in Materials & methods section. The Bayesian framework is both straightforward and efficient in propagating error through the parameters. We used uninformative priors for all parameters. Since the maximum growth rate is a life history trait expected to be stable between populations of the same species, we used an uninformative hierarchical prior for this parameter. We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, and discarded the first 2,000 samples as burn-in. We checked the chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992). The models were fitted using NIMBLE [@deValpine2017] run from R [@RCT2022]. The values **X[Y; Z]** reported in Results section are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions. We preferred the median to the mean because it is more robust to skewed distributions.

@@

\setstretch{1}
```{r vital, fig.cap = "(ref:vital)"}
knitr::include_graphics("../Output/plot_paper_3.png")
```
\setstretch{2}

@@

When there was no culling pressure, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), namely 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for GB and FR, respectively. These values corresponded to population increases of 45% and 52% per year, respectively. Assuming a maximum adult survival rate of 0.85 [0.7; 1] for both populations, we derived maximum recruitment rates of 0.60 [0.42; 0.78] and 0.68 [0.36; 1.06] for GB and FR, respectively.
